# Fast food entree data

* Data scraped from [fastfoodnutrition.com](https://fastfoodnutrition.org/mcdonalds/chart) 
* Please notice that I really only took entrees - feel free to select ALL food, sides, drinks, desserts, etc.

Code to webscrape seen below.
```{r libraries}
## Libraries
library(tidyverse)
library(rvest)
```


Our core url for example is  [https://fastfoodnutrition.org/mcdonalds/chart](https://fastfoodnutrition.org/mcdonalds/chart). The only portion that changes between fast food joints is the name.

Goal is to selection specific fast food restaurant and then bind all the various tables into one final dataframe.


```{r web_scrape_function}
food_scrape <- function(restaurant, tbl_sel){
    url <- glue::glue("https://fastfoodnutrition.org/{restaurant}/chart")
    url %>% 
        read_html() %>% 
        html_table() %>% 
        .[tbl_sel] %>% 
        bind_rows() %>% 
        select(-X16) %>%
        set_names(nm = c("item", "calories", "cal_fat", "total_fat", "sat_fat", "trans_fat",
                         "cholesterol", "sodium", "total_carb", "fiber", "sugar", "protein",
                         "vit_a", "vit_c", "calcium")) %>% 
        mutate(restaurant = str_replace(restaurant, "-", " "),
               restaurant = str_to_title(restaurant)) # save restaurant name
} 
```

# Combine datasets and save

```{r}
mcd_df <- food_scrape("mcdonalds", c(1,2,3,9,19))
cfa_df <- food_scrape("chick-fil-a", c(1,2,8,13))
sonic_df <- food_scrape("sonic", c(1,2,17,18,20))
arbys_df <- food_scrape("arbys", c(1:4,8))
bk_df <- food_scrape("burger-king", c(1:3,7,11:12))
dq_df <- food_scrape("dairy-queen", c(5, 7, 12, 25, 27))
sub_df <- food_scrape("subway", c(1,2,3,4,5,6,7,8,9))
taco_df <- food_scrape("taco-bell", c(1,2,3,4,5,15,18,19,20,22,23,24))

final_df <- bind_rows(mcd_df, cfa_df, sonic_df, arbys_df, bk_df, dq_df, sub_df, taco_df) %>% 
    select(restaurant, everything()) %>% 
    mutate(salad = case_when(str_detect(item, "salad") ~ "Salad",
                             TRUE ~ "Other"))

final_df %>% write_csv("fastfood_calories.csv")
```
